### 2024-08-20 manuscript/supplementary.tex

\documentclass{article}
\usepackage{graphicx}
\usepackage{soul,xcolor}

\newcommand\coordinator[1]{\begin{flushleft}\small\textit{#1}\end{flushleft}}

\title{Proceedings of the OHBM Hackathon 2023 -- Supplementary Material}
\begin{document}
\maketitle

\framebox[\linewidth]{\parbox{\textwidth}{
Bedtime Story for Hackers \\
"The Chat in the Hack Comes Back" \\
Variation of a theme from Dr. Suess's "The Cat in the Hat Comes Back" \\
by Tonya White \\
\\
There is no time for play, there is no time for fun, \\
There is no time for games, there is work to be done. \\
When my PI went back to her humble abode, \\
She said somebody has to finish this code. \\
Somebody, somebody has to you see. \\
So she looked at two somebodies, the postdoc and me. \\
\\
So we looked at each other, and what could we do. \\
Hacking is fun, and much better with two. \\
But it's even more fun, with three, four or eight, \\
Make it a group, a hackathon would be great. \\
\\
So I sent out the message to all on discord, \\
Come now, come quick, come all in accord, \\
There shall be coffee, there will be space, \\
There will be Internet, and a wonderful place. \\

With heads full of brains, the room started to fill,  \\
The projects were shared, the coders were chill. \\
The coffee was hot, the projects were cool. \\
The computers were buzzing with their underlying boel… \\
…ian logic.  \\
\\
So there we all were, all hacking so fast, \\
But who should walk by, but the chat in the hack. \\
Your coding's in vain, the chat in the hack said, \\
With chatGPT, your coding is dead! \\
\\
Not so, not so! We cried in accord, \\
We're creating new things, new things galore \\
Your ChatGPT takes only what's there, \\
It can't create things, just poof from thin air. \\
\\
So we presented our projects, as we all sat around, \\
All remained still, still with no other sounds, \\
Except the sound of sips, the caffeinated type. \\
As we all looked in awe, seeing brain power and might. \\
\\
After all the talks, and all the applause, \\
The room remained still, at least for a brief pause, \\
Then the chat in the hack laid down his hat, \\
I surrender he said, I take my words back. \\
Your hackathon is great, even though I was late, \\
Next year I'll be here, I'll make it a date. \\
}}

\framebox[\linewidth]{\parbox{\textwidth}{
Untitled \\
James D Kent \\
\\
Dip your toes into neurosynth compose, \\
Cause we're pros we got immaculate workflows. \\
It's the platform you'll be glad you chose \\
Shameless plug https://compose.neurosynth.org/ \\
Now sit back and listen to my prose \\
Coordinate based regression is my current obsession, \\
Yifan’s hunting bugs with relentless aggression. \\
she does it with a smile and a calming expression. \\
Optimizing functions with her matrices hessian. \\
We're not done, we got image based analyses. \\
Propelling science to interstellar galaxies. \\
Lea’s been working on fixing technicalities, \\
Bringing science fiction into our very own realities. \\
Lest I forget, there is Kendra with Neurovault \\
The data processed just crumbles with her assault \\
breaking down abstracts like grinding the table salt \\
her flow is like lava, ideas solid like rock basalt \\
And there's more, we made a tutorial. \\
Here in Montreal, a region quite boreal \\
We be spreading knowledge like we bacterial. \\
If you think that's cool come check our material. \\
So, dip your toes into neurosynth compose \\
It's the platform you'll be glad you chose. \\
}}

\framebox[\linewidth]{\parbox{\textwidth}{
Untitled \\
François Paugam \\
\\
Now this is a story \\
All about how \\
My life got flipped \\
Turned upside down \\
And I’d like to take a minute \\
So just listen \\
I’ll tell you how I became contributor \\
of a package called nilearn \\
\\
Look \\
If you had \\
One issue \\
Or one pull request \\
To refactor every test ever created \\
In one branch \\
Would you merge it? \\
Or just let it slip? \\
\\
His code is ready \\
Nice tweaks, lines are pretty \\
There’s a commit on his branch already \\
Code spaghetti \\
\\
He’s nervous, but on Github \\
He looks calm and ready \\
To fix bugs \\
But he keeps on forgettin’ \\
\\
What he wrote down \\
All the tests go so bad \\
He opens a PR \\
But the merge won’t come out \\
He’s chocking, how? \\
Everybody’s joking now \\
The clock’s run out  \\
flake8 failed, blaow! \\
\\
Don’t git push \\
Cause I ain’t run flake8 \\
I’m trying \\
Not to fail the tests \\
It’s like spaghetti sometimes \\
It makes me wonder how I keep from getting errors \\
Don’t need no GPT hallucinating \\
I get my code hints from Hao-Ting \\
\\
Rhyming or coding \\
It’s all the same \\
Find fine lines online \\
Copy-paste them \\
Have an issue with that? \\
You know who to git-blame \\
}}

\end{document}