### 2024-12-13 manuscript/affiliations.tex
enh: more adjustnemnts to text

\affil[1]{Division of Experimental Psychology and Neuropsychology, Department of Education and Psychology, Freie Universität Berlin}
\affil[2]{Department of Psychology, Center for Perceptual Systems, The University of Texas at Austin}
\affil[3]{Centre for Sleep and Cognition & Centre for Translational Magnetic Resonance Research, Yong Loo Lin School of Medicine, National University of Singapore, Singapore}
\affil[4]{Department of Psychiatry, The University of Melbourne and Melbourne Health, Victoria, Australia}
\affil[5]{Shanghai Mental Health Center, School of Medicine, Shanghai Jiao Tong University, 600 S. Wanping Rd., Shanghai, 200030}
\affil[6]{Department of Psychology, Faculty of Arts and Social Science, University of Zurich, Zurich, ZH, Switzerland}
\affil[7]{School of Electrical and Computer Engineering, Georgia Institute of Technology, Atlanta, Georgia, United States}
\affil[8]{Department of Cognitive Neuroscience, Faculty of Psychology and Neuroscience, Maastricht University, Maastricht, The Netherlands}
\affil[9]{School of Physics, The University of Sydney, Sydney, Australia}
\affil[10]{Big Data Institute, University of Oxford, UK}
\affil[11]{Univ Rennes, Inria, CNRS, Inserm, Rennes, France}
\affil[12]{Department of Radiology, Lausanne University Hospital and University of Lausanne, Lausanne, VD, Switzerland.}
\affil[13]{Robarts Research Institute, Western University, London, Ontario, Canada}
\affil[14]{School of Psychological Sciences, College of Engineering, Science and Environment, University of Newcastle, Newcastle, NSW, Australia}
\affil[15]{Hunter Medical Research Institute, Newcastle, NSW, Australia}
\affil[16]{Speech and Hearing Bioscience and Technology, Harvard University, USA}
\affil[17]{Campbell Family Mental Health Research Institute, Centre for Addiction and Mental Health, Toronto, ON, Canada}
\affil[18]{Northwestern University, Evanston, USA}
\affil[19]{Department of Bioengineering, McGill University, Montreal, QC, Canada}
\affil[20]{Paris Brain Institute}
\affil[21]{Department of Radiography, Faculty of Health and Social Sciences, Western Norway University of Applied Sciences, Bergen, Norway}
\affil[22]{Centre de recherche de l'Institut universitaire de gériatrie de Montréal, Montréal, QC, Canada}
\affil[23]{Brain Innovation, Maastricht, The Netherlands}
\affil[24]{National Institutes of Health, Bethesda, MD, USA}
\affil[25]{Department of Biomedical Engineering, Vanderbilt University}
\affil[26]{Danish Research Centre for Magnetic Resonance, Centre for Functional and Diagnostic Imaging and Research, Copenhagen University Hospital - Amager and Hvidovre, Copenhagen, Denmark}
\affil[27]{Melbourne Neuropsychiatry Centre, Department of Psychiatry, The University of Melbourne and Melbourne Health, Victoria, Australia}
\affil[28]{Neuro-X Institute, École polytechnique fédérale de Lausanne, Geneva, Switzerland}
\affil[29]{Department of Radiology and Medical Informatics (DRIM), Faculty of Medicine, University of Geneva, Geneva, Switzerland}
\affil[30]{Department of Cognitive Neuroscience, Faculty of Psychology and Neuroscience, Maastricht University, Maastricht, The Netherlands}
\affil[31]{Big Data Institute, Li Ka Shing Centre for Health Information and Discovery, Old Road Campus, Oxford, OX3 7LF, UK}
\affil[32]{Fetal Neonatal Neuroimaging and Developmental Science Center, Boston Children’s Hospital, Harvard Medical School, Boston, MA, United States}
\affil[33]{Organization for Human Brain Mapping, USA}