### 2025-04-13 manuscript/affiliations.tex
enh: authors & affiliations

\affil[1]{Division of Experimental Psychology and Neuropsychology, Department of Education and Psychology, Freie Universität Berlin}
\affil[2]{OHBM Open Science Special Interest Group}
\affil[3]{Department of Psychology, Center for Perceptual Systems, The University of Texas at Austin, Austin, Texas, United States}
\affil[4]{Fetal Neonatal Neuroimaging and Developmental Science Center,Boston Children’s Hospital, Harvard Medical School, Boston,United States}
\affil[5]{Department of Psychology, Faculty of Arts and Social Science, University of Zurich, Zurich, ZH, Switzerland}
\affil[6]{Shanghai Mental Health Center, School of Medicine, Shanghai, Jiao Tong University, Shanghai, China}
\affil[7]{Melbourne Neuropsychiatry Centre, Department of Psychiatry, The University of Melbourne and Melbourne Health, Victoria, Australia}
\affil[8]{Centre for Sleep and Cognition Centre for Translational, Magnetic Resonance Research, Yong Loo Lin School of Medicine,National University of Singapore, Singapore}
\affil[9]{Department of Biomedical Engineering, The University of Melbourne, Victoria, Australia}
\affil[10]{Department of Psychiatry, The University of Melbourne and Melbourne Health, Victoria, Australia}
\affil[11]{School of Electrical and Computer Engineering, Georgia Institute of Technology, Atlanta, Georgia, United States}
\affil[12]{Organization for Human Brain Mapping, Roseville, United States}
\affil[13]{National Institute of Mental Health, Bethesda, MD, USA}
\affil[14]{School of Physics, The University of Sydney, Sydney, Australia}
\affil[15]{McGill, Montreal Neurological Institute - Hospital, Montreal, Quebec, Canada}
\affil[16]{Campbell Family Mental Health Research Institute, Centre for Addiction and Mental Health, Toronto, ON, Canada}
\affil[17]{Univ Rennes, Inria, CNRS, Inserm, Rennes, France}
\affil[18]{Child Mind Institute, New York, NY, USA}
\affil[19]{Gina Cody School of Engineering and Computer Science, Concordia University, Montreal, QC, Canada}
\affil[20]{Data Science and Sharing Team (DSST), Intramural Research Program, NIMH, Bethesda, USA}
\affil[21]{Department of Biomedical Engineering, Vanderbilt University}
\affil[22]{Big Data Institute, Li Ka Shing Centre for Health Information and Discovery, Old Road Campus, Oxford, UK}
\affil[23]{Cognitive Neuroscience Department, Faculty of Psychology and Neuroscience, Maastricht University, Maastricht, Netherlands}
\affil[24]{School of Psychological Sciences, College of Engineering, Science and Environment, University of Newcastle, Newcastle, NSW, Australia}
\affil[25]{Hunter Medical Research Institute, Newcastle, NSW, Australia}
\affil[26]{Paris Brain Institute}
\affil[27]{Robarts Research Institute, Western University, London, Ontario, Canada}
\affil[28]{Northwestern University, Evanston, USA}
\affil[29]{Department of Cognitive Neuroscience, Faculty of Psychology and Neuroscience, Maastricht University, Maastricht, The Netherlands}
\affil[30]{Department of Radiography, Faculty of Health and Social Sciences, Western Norway University of Applied Sciences, Bergen, Norway}
\affil[31]{Mohn Medical Imaging and Visualization Centre (MMIV), Bergen, Norway}
\affil[32]{NeuroDataSciene - ORIGAMI lab, Montreal Neurological Institute-Hospital, McGill University, Montreal, Canada}
\affil[33]{Program in Speech and Hearing Bioscience and Technology, Division of Medical Sciences, Harvard University, Cambridge, MA, USA}
\affil[34]{Department of Brain and Cognitive Sciences, Massachusetts Institute of Technology, Cambridge, MA, USA}
\affil[35]{Speech and Hearing Bioscience and Technology, Harvard University, USA}
\affil[36]{Department of Bioengineering, McGill University, Montreal, QC, Canada}
\affil[37]{Neuro-X Institute, École polytechnique fédérale de Lausanne, Geneva, Switzerland}
\affil[38]{physiopy.github.io}
\affil[39]{Department of Psychology, University of Montreal, Canada}
\affil[40]{Centre de recherche de l'Institut universitaire de gériatrie de Montréal, Montréal, QC, Canada}
\affil[41]{Department of Radiology, Lausanne University Hospital and University of Lausanne, Lausanne, VD, Switzerland.}
\affil[42]{Danish Research Centre for Magnetic Resonance, Centre for Functional and Diagnostic Imaging and Research, Copenhagen University Hospital - Amager and Hvidovre, Copenhagen, Denmark}
\affil[43]{Medical Physics Department, University Hospital Southampton NHS Foundation Trust, Southampton, England, United Kingdom}
\affil[44]{School of Clinical and Experimental Sciences, Faculty of Medicine, University of Southampton, Southampton, England, United Kingdom}
\affil[45]{Department of Cognitive Science and Artificial Intelligence, Tilburg School of Humanities and Digital Sciences, Tilburg University, Tilburg, The Netherlands}
\affil[46]{Donders Centre for Cognitive Neuroimaging, Donders Institute for Brain Cognition and Behaviour, Radboud University, Nijmegen, The Netherlands}